---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
#install.packages('dplyr')
#install.packages('Seurat', repos = "http://cran.us.r-project.org")
#install.packages('patchwork', repos = "http://cran.us.r-project.org")


```


```{r}
library(dplyr)
# retrieve subset data
counts_exon <- read.csv("mouse_VISp_2018_exon-matrix.csv")
counts_intron <- read.csv("mouse_VISp_2018_intron-matrix.csv")
sample_info <- read.csv("mouse_VISp_2018_samples-columns.csv")
```

# filter for only GABAergic cells


```{r}
# need to make first row as index to for next functions to work
rownames(counts_exon) <- counts_exon[,1]
counts_exon <- counts_exon[,-1]
```


# SEURAT

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
```

```{r}
# create seurat object
exon_seurat <- CreateSeuratObject(counts = counts_exon, project = "Tasic_exon", min.cells = 3, min.features = 200)
```

1. Selection and filtration of cells based on QC metrics, data normalization and scaling, and the detection of highly variable features
```{r}
# Getting rid of mitochondrial genes
exon_seurat[["percent.mt"]] <- PercentageFeatureSet(exon_seurat, pattern = "^MT-") # calculates the percentage of counts originating from mitochondria features
head(exon_seurat@meta.data)
```

```{r}
VlnPlot(exon_seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, sort = TRUE)

plot1 <- FeatureScatter(exon_seurat, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(exon_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```
```{r}
exon_seurat <- subset(exon_seurat, subset = nFeature_RNA > 2000 & nFeature_RNA < 15000 & percent.mt < 5) 

```


2. Normalizing data after removing unwanted cells from the dataset
```{r}
#  global-scaling normalization method 
exon_seurat <- NormalizeData(exon_seurat, normalization.method = "LogNormalize", scale.factor = 10000)
```
3. Identifying highly variable features/genes
```{r fig.height = 3, fig.width = 9}
# First plot mean-variance of each gene across all cells
#exon_seurat1 <- FindVariableFeatures(exon_seurat,selection.method = "vst", nfeatures = 2000 ) 
exon_seurat <- FindVariableFeatures(exon_seurat,selection.method = "mean.var.plot") 
# how to know if 2000 is good? look at variance of all genes and see when it apers off. Calculate variance across geen. using seurat object

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(exon_seurat)
#plot3 <- VariableFeaturePlot(exon_seurat1)
#plot2 <- LabelPoints(plot = plot1, points = top1000, repel = TRUE, xnudge = 0, ynudge = 0) 
#plot2
plot1


```


4. Selecting highly variable genes
```{r}
# Identify the 1000 most highly variable genes - play around later
top1000 <- head(VariableFeatures(exon_seurat), 1000)
```



5. Scale genes 
```{r}
all.genes <- rownames(exon_seurat)
exon_seurat <- ScaleData(exon_seurat, features = all.genes)
```

5. Linear dimensionality reduction for HVGs
```{r}
exon_seurat <- RunPCA(exon_seurat, features = VariableFeatures(object = exon_seurat), npcs = 100)
```

# Seurat provides several useful ways of visualizing both cells and features that define the PCA, including VizDimReduction(), DimPlot(), and DimHeatmap()
```{r fig.height = 4, fig.width = 8}
# Examine and visualize PCA results a few different ways
print(exon_seurat[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(exon_seurat, dims = 1:2, reduction = "pca")

DimPlot(exon_seurat, reduction = "pca")
```

```{r}
DimHeatmap(exon_seurat, dims = 1, cells = 500, balanced = TRUE) # only PCA 1

DimHeatmap(exon_seurat, dims = 1:10, cells = 500, balanced = TRUE)
# imporatnacce of the variance gene to the PC
```
```{r}
ElbowPlot(exon_seurat, ndims = 100)
```





